Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-06-05 paper/long_version.tex
Auto Update from local Sharelatex

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }
\usepackage{wrapfig}        % added by seth for placing images in-line with text

@@

\author{
  James King, Seth Russell, Tellen D. Bennett, Debashis Ghosh\\
  Data Science to Patient Value (D2V)\\
  School of Medicine\\
  University of Colorado Anschutz Medical Campus\\
  Aurora, CO\\
  \texttt{\{james.king, seth.russell, tell.bennett, debashis.ghosh\}@ucdenver.edu}
  }

@@

% this tool simplfies data sharing, simplifies data exploration within organizations
\begin{abstract}
There are many barriers to data access and data sharing, especially in the domain of machine learning on health care data. Legal constraints such as HIPAA protect patient privacy but slow access to data and limit reproducibility. We provide a description of an end-to-end system called \emph{Kung Faux Pandas} for easily generating de-identified or synthetic data which is statistically similar to given real data but lacks sensitive information. This system focuses on data synthesis and de-identification narrowed to a specific research question to allow for self-service data access without the complexities required to generate an entire population of data that is not needed for a given research project. Kung Faux Pandas is an open source publicly available\footnote{\url{https://github.com/CUD2V/kungfauxpandas}} system that lowers barriers to HIPAA- and GDPR-compliant data sharing for enabling reproducibility and other purposes.
\end{abstract}

@@

Recent work has clarified the definition and importance of reproducibility and a closely related concept, replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result."\cite{leek_opinion_2015,peng_reproducible_2006} Others have used the terms in a reverse fashion.\cite{drummond_replicability_2009} Despite the semantic differences, there is broad agreement that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing lists the key factors in reproducibility as available data, input parameters, documentation, software code, and an environment capable of running the provided software code. These items are difficult to completely convey in a traditional research paper \cite{schwab_making_2000}.

@@

Many domains in which ML investigators work have legal barriers that make it difficult to enable independent verification of ML models which frequently require enormous amounts of training data. As an example, in the health care domain in the United States, a key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about security and what can and cannot be done with protected health information \cite{hippaviol}. While there are other domain specific data privacy laws, there are also some broad rules that affect entire countries such as the European Union's General Data Protection Regulation (GDPR). The GDPR puts in place new rules about data ownership and control in an attempt to give individuals more control over their information.

@@

As the data generated through KFP has no protected component, it can be shared along with code or methods to improve reproducibility as well as aid in replicability. To further the aim of reproducibility we have provided our code and a sample of a randomly generated data set for testing purposes through our GitHub repository in a range of formats: Python environment files (conda and pip) to allow others to start with the same software libraries and specific versions that we used, a Docker container with code and sample data installed and already configured, and all source code along with version history so others can follow our development process.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{prototype_architecture}
  \caption{System Architecture.}
  \label{fig:architecture}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}

@@

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible ML. Through providing ad-hoc faux-data, ML researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing minimum dataset size issues, and evaluation of this system in a ML workflow.

@@

Another area that is open for further research is that of privacy protection through use of minimum dataset size. In the field of education, statistical methods for calculating sample size based on probability of detecting a specific effect size are commonly used \cite{naep_2009}. However, it is not clear that these methods are appropriate for use ML in general or the specific domain of health care. Furthermore, effect size and probability to detect specific effects may vary by use case. Thus, in this work, no attempt has been made to constrain data set sizes or results returned. KFP could be modified to return no results if a minimum threshold is not met based on the risk profile of the data owner or, alternatively, data boosting techniques in combination with data synthesis could be used.

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.

% include some of this? 
% Lastly, the Synthetic dataset Generation Framework (SGF) by Bindschaedler et al. \cite{Bindschaedler2017} "generates differentially private synthetic data" using a novel criterion called plausible deniability that is applied to a dataset before it is deemed safe to release.  In the case of SGF, several models are built with both real and synthetic data and compared against a gold standard. Their author's analysis showed that the models from synthetic data are close to the same accuracy as models built with real data.

>>>>

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }

@@

\author{
  James King, Seth Russell, Tellen D. Bennett, Debashis Ghosh\\
  Data Science to Patient Value (D2V), School of Medicine\\
  University of Colorado Anschutz Medical Campus, Aurora, CO\\
  \texttt{\{james.king, seth.russell, tell.bennett, debashis.ghosh\}@ucdenver.edu}
  }

@@

\begin{abstract}
There are many barriers to data access and data sharing, especially in the domain of machine learning on health care data. Legal constraints such as HIPAA protect patient privacy but slow access to data and limit reproducibility. We provide a description of an end-to-end system called \emph{Kung Faux Pandas} for easily generating de-identified or synthetic data which is statistically similar to real data but lacks sensitive information. This system focuses on data synthesis and de-identification narrowed to a specific research question to allow for self-service data access without the complexities required to generate an entire population of data that is not needed for a given research project. Kung Faux Pandas is an open source publicly available\footnote{\url{https://github.com/CUD2V/kungfauxpandas}} system that lowers barriers to HIPAA- and GDPR-compliant data sharing for enabling reproducibility and other purposes.
\end{abstract}

@@

Recent work has clarified the definition and importance of reproducibility and a closely related concept, replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result."\cite{leek_opinion_2015,peng_reproducible_2006} Others have used the terms in a reverse fashion\cite{drummond_replicability_2009}. Despite the semantic differences, there is broad agreement that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing lists the key factors in reproducibility as available data, input parameters, documentation, software code, and an environment capable of running the provided software code. These items are difficult to completely convey in a traditional research paper \cite{schwab_making_2000}.

@@

Many domains in which ML investigators work have legal barriers that make it difficult to enable independent verification of ML models which frequently require enormous amounts of training data. As an example, in the health care domain in the United States, a key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about data security and what can and cannot be done with protected health information \cite{hippaviol}. While there are other domain specific data privacy laws, there are also some broad rules that affect entire countries such as the European Union's General Data Protection Regulation (GDPR). The GDPR puts in place new rules about data ownership and control in an attempt to give individuals more control over their information.

@@

As the data generated through KFP has no protected component, it can be shared along with code or methods to improve reproducibility as well as aid in replicability. To further the aim of reproducibility we have provided our code and a sample of a randomly generated dataset for testing purposes through our GitHub repository in a range of formats: Python environment files (conda and pip) to allow others to start with the same software libraries and specific versions that we used, a Docker container with code and sample data installed and already configured, and all source code along with version history so others can follow our development process.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{prototype_architecture}
  \caption{System Architecture.}
  \label{fig:architecture}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}

@@

KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{figure}

@@

KFP provides a unique contribution to the space of reproducible ML. Through providing ad-hoc faux-data, ML researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing institutional specific privacy issues, and evaluation of this system in a ML workflow.

@@

Another area that is open for further research is that of domain and institution specific data privacy needs. In the field of education, statistical methods for calculating sample size based on probability of detecting a specific effect size are commonly used \cite{naep_2009} as the key method for privacy protection. However, it is not clear that these methods are appropriate for use in ML in general or domains other than education. Thus, in this work, no attempt has been made to constrain data set sizes nor provide additional proofs of data protection. KFP could be modified to return no results if a minimum threshold is not met or, alternatively, data boosting techniques in combination with data synthesis could be used based on the risk profile of the data owner(s).

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.
% include some of this? 
% Lastly, the Synthetic dataset Generation Framework (SGF) by Bindschaedler et al. \cite{Bindschaedler2017} "generates differentially private synthetic data" using a novel criterion called plausible deniability that is applied to a dataset before it is deemed safe to release.  In the case of SGF, several models are built with both real and synthetic data and compared against a gold standard. Their author's analysis showed that the models from synthetic data are close to the same accuracy as models built with real data.

### 2018-06-05 paper/seth.tex
Auto Update from local Sharelatex

>>>>

 % Some information about minimum group size
 %* Minimum group/result set size depends on several factors such as re-identification risk, number of attributes available (see picture at end)
 %* Individual states have adopted minimum group size reporting rules, with the minimum number of students ranging from 5 to 30 and a modal category of 10 (used by 39 states) https://nces.ed.gov/pubs2011/2011603.pdf
 %* 62 https://nces.ed.gov/nationsreportcard/tdw/analysis/summary_rules_minimum.aspx
 %* 30; describes techniques to determine size - https://nces.ed.gov/pubs2017/2017147.pdf

 % removed since we never got SGF working
 \begin{table}
  \caption{Kung Faux Pandas data synthesis methods}
  \label{included-methods}
  \centering
  \begin{tabular}{p{12em} l p{14em}}
    \toprule
    Method                                 & Source                    & Notes \\
    \midrule
    Kernel Density Estimator               & scipy.stats.gaussian\_kde & Inter-related ordinal and ratio data\\
    DataSynthesizer                        & Ping et al.               & Bayesian modeling of column relationships\\
    Synthetic dataset Generation Framework & Bindschaedler et al.      & "Plausible Deniability" probabilistic modeling\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}

### 2018-06-05 paper/short_version.tex
Auto Update from local Sharelatex

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }

@@

% where to put this text? "Unit testing is important but often overlooked - unit tests can show how a researcher validated their code was giving expected results which can often lead to insights not normally communicated in a results section of a journal article."
One way around the difficulties of source data access is to separate the dependency between data and analysis in such a way that analysis can be performed without the analyst having access to the source data. Such a process can be possible if a sufficiently detailed description of the sensitive data is provided. Appropriate models can be developed and initially validated before either being handed off to a data custodian who then runs the processes on protected data and returns the results to the researcher or access to the protected data has been granted directly to the researcher. Solutions of this form have been possible for some time, but are awkward in practice because activities such as data cleaning, exploration, and model building are interactive and iterative processes that require more than just descriptions of data.

@@

Ideally a ML investigator should develop, test, and validate models using all available data. This is most easily accomplished when they have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} shows the novel architecture of KFP: data access through an intermediary that can synthesize data on-demand.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{prototype_architecture}
  \caption{System Architecture.}
  \label{fig:architecture}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}

Figure~\ref{fig:synthesis_process} shows the automated faux-data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, ML developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux-data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.

KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{figure}

@@

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.
% include some of this? 
% Lastly, the Synthetic dataset Generation Framework (SGF) by Bindschaedler et al. \cite{Bindschaedler2017} "generates differentially private synthetic data" using a novel criterion called plausible deniability that is applied to a dataset before it is deemed safe to release.  In the case of SGF, several models are built with both real and synthetic data and compared against a gold standard. Their author's analysis showed that the models from synthetic data are close to the same accuracy as models built with real data.

>>>>

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }
\usepackage{wrapfig}        % added by seth for placing images in-line with text

@@

One way around the difficulties of source data access is to separate the dependency between data and analysis in such a way that analysis can be performed without the analyst having access to the source data. Such a process can be possible if a sufficiently detailed description of the sensitive data is provided. Appropriate models can be developed and initially validated before either being handed off to a data custodian who then runs the processes on protected data and returns the results to the researcher or access to the protected data has been granted directly to the researcher. Solutions of this form have been possible for some time, but are awkward in practice because activities such as data cleaning, exploration, and model building are interactive and iterative processes that require more than just descriptions of data.

@@

Ideally a ML investigator should develop, test, and validate models using all available data. This is most easily accomplished when they have unfettered access to all data in the problem domain they are working in. The novel architecture of KFP is that direct data access is provided through an intermediary that can synthesize data on-demand.

\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}

Figure~\ref{fig:synthesis_process} shows the automated faux-data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, ML developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux-data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.
\begin{wrapfigure}[16]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.
